### 2020-01-31 misc/meeting_notes.Rmd
Added meeting notes

---
output:
  pdf_document: default
  html_document: default
---

## Meeting notes
* Company helps appeal assessment values for residential and commercial 
* Both residential and commercial are important to the model, although the majority of company's profits come from corporate clients
* Tax classes (classified by government, with 01 residential, 05 light industrial, 06 commercial being the importnat classes)
* Some condos + commercial complexes have split classes
* Mill rate is used interchangeably with tax rate, and is unique to a  specific code (called mill rate because its over 1000)
* eg: “12.3* assessment /1000” mill rate is 12.3
* Three things he is interested in primarily:
     1. Predicting mill rate (published in April but assessments published in January) which changes every year, during the blank period from jan-april
     2. Predicting next years assessment values (once mill rates are known, for the period between May-Dec)
     3. Predicting final tax that will be owed
* Their service is used to help with tax budgeting
* Sometimes values are missing, but we still need to predict these things given what we do have available for a given property
* Municipal government has budget, and determines mill rates for each tax code in order to raise the budget amount (from other funds also, but estimated about 90% of fund raised by property taxes)
* Mentioned it would be interesting to quantify relationship between mill rates and assessment values, but accurate prediction is priority
* Roll number uniquely identifies properties across datasets (see handout for the form of this identifier)
* 2016-2020 data is available (2015?)
* Vancouver municipality is a special case when it comes to tax coding - this should be treated as such and is not comparable to other municipalities
* Sales data is available (they bought this from government); each row is a sales transaction (filter to market sales only, to leave name transfers out of it)
* Currently if property assessments go up 10%, they are assuming mill rate goes down 6-7% due to inflation and larger budgets and predicting like this (it's not working well)
* Assumes large chunk of variability in mill rate is due to assessment (and not just municipal budget amount)
* Table handout is for changes in assessment values (are they averages? he's not sure but we could check)
* Exemption exists but is maybe not that important
* Whatever information is used for assessment by the government is also available to us