### 2020-02-25 src/tax2016-2018.Rmd
filtered tax for municipalities in metro vancouver from 2016-2018 and computed percentage change of tax from 2017-2018.

---
title: "budget"
author: "Xuechun Lu"
date: "25/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
```

```{r}
schedule2016<-read_excel(here("data/tax", "schedule706_2016.xls"))
schedule2017<-read_excel(here("data/tax", "schedule706_2017.xls"))
schedule2018<-read_excel(here("data/tax", "schedule706_2018.xls"))

colnames(schedule2016) <- schedule2016[1,]
schedule2016<-schedule2016[-c(1),]
colnames(schedule2017) <- schedule2017[1,]
schedule2017<-schedule2017[-c(1),]
colnames(schedule2018) <- schedule2018[1,]
schedule2018<-schedule2018[-c(1),]

municipality.list = c(
  "Burnaby", 
  "Coquitlam", 
  "Delta", 
  "Langley", 
  #"Langley - Township",
  "Maple Ridge",
  #"Maple Ridge Rural", 
  "North Vancouver",
  #"North Vancouver - Dist",
  "Pitt Meadows", 
  "Port Coquitlam", 
  "Port Moody", 
  "Richmond", 
  "Surrey", 
  "Vancouver", 
  "White Rock", 
  "West Vancouver", 
  "Bowen Island", 
  "Anmore", 
  "Belcarra",
  "Lions Bay")

tax.2016<-schedule2016 %>% filter(Municipalities %in% municipality.list) %>% 
  select("Municipalities","Type","Current Year Taxes Collected")
tax.2017<-schedule2017 %>% filter(Municipalities %in% municipality.list) %>% 
  select("Municipalities","Type","Current Year Taxes Collected")
tax.2018<-schedule2018 %>% filter(Municipalities %in% municipality.list) %>% 
  select("Municipalities","Type","Current Year Taxes Collected")

# rename
tax.2016[7,1]<-"Langley - City"
tax.2016[8,1]<-"Langley - Township"
tax.2016[11,1]<-"North Vancouver - City"
tax.2016[12,1]<-"North Vancouver - Dist"

tax.2017[7,1]<-"Langley - City"
tax.2017[8,1]<-"Langley - Township"
tax.2017[11,1]<-"North Vancouver - City"
tax.2017[12,1]<-"North Vancouver - Dist"

tax.2018[7,1]<-"Langley - City"
tax.2018[8,1]<-"Langley - Township"
tax.2018[11,1]<-"North Vancouver - City"
tax.2018[12,1]<-"North Vancouver - Dist"

tax.2016<-tax.2016 %>% select("Municipalities","Current Year Taxes Collected")
tax.2017<-tax.2017 %>% select("Municipalities","Current Year Taxes Collected")
tax.2018<-tax.2018 %>% select("Municipalities","Current Year Taxes Collected")

names(tax.2016)[2]<-paste("2016")
names(tax.2017)[2]<-paste("2017")
names(tax.2018)[2]<-paste("2018")
tax.full<-tax.2016 %>% left_join(tax.2017, by = "Municipalities") %>% 
   left_join(tax.2018, by = "Municipalities")

# convert chr to numeric
tax.full[,2]<-as.numeric(tax.full$`2016`)
tax.full[,3]<-as.numeric(tax.full$`2017`)
tax.full[,4]<-as.numeric(tax.full$`2018`)

# compute percentage change

Municipalities<-tax.full$Municipalities
pct_2017<-(tax.full$`2017`-tax.full$`2016`)/tax.full$`2016`*100
pct_2018<-tax.pct[,3]<-(tax.full$`2018`-tax.full$`2017`)/tax.full$`2017`*100
tax.pct<-data.frame(Municipalities,pct_2017,pct_2018)
```

Note: 
- tax.full contains tax from 2016-2018 for 20 municipalities
- tax.pct contains percentage change of tax from 2017-2018.