### 2020-02-28 src/EDA&mode_fitting.Rmd
cleanup and add graphs

---
title: "EDA"
author: "Xuechun Lu, Yuting Wen, Peter Han, Yuetong Liu"
date: "25/02/2020"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
libraries
```{r, message=F, warning=F}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
library(readxl)
```

```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

1. data processing
```{r, warning=F, message=F}
# all municipalities
municipality.list = c(
  "Burnaby", 
  "Coquitlam", 
  "Delta", 
  "Langley - City", 
  "Langley - Township",
  "Maple Ridge",
  "Maple Ridge Rural", 
  "North Vancouver - City",
  "North Vancouver - Dist",
  "Pitt Meadows", 
  "Port Coquitlam", 
  "Port Moody", 
  "Richmond", 
  "Surrey", 
  "Vancouver", 
  "White Rock", 
  "West Vancouver", 
  "Bowen Island", 
  "Anmore", 
  "Belcarra",
  "Lions Bay")

# budget and assessment data import and process
tax_pct<-read_csv(here("data", "tax_pct.csv"))
assessment_pct<-read_csv(here("data", "assessment_pct.csv"))
# omit 2016
assessment_pct<-na.omit(assessment_pct)
assessment_pct<-assessment_pct[,-c(1)]
names(assessment_pct)[2]<-paste("Municipalities")
tax_pct<-tax_pct[,-c(1)]
#head(tax_pct)

assessment_aggregate <- read.csv(here("data","assessment_aggregate.csv"))
```
Note: tax_pct doesn't have Maple Ridge Rural. We decide to treat Maple Ridge Rural and Maple Ridge with the same tax.

2019 tax pct is missing, we imputed 2019 tax by the average of previous tax pct
```{r}
tax_pct[,4]<-(tax_pct$pct_2017+tax_pct$pct_2018)/2
#tax_pct
names(tax_pct)[4]<-paste("pct_2019")
tax_pct<-rbind(tax_pct,c("Maple Ridge Rural", 6.195736, 5.6949567,5.945346))

#head(tax_pct)
```

aggregate tax and assessment
```{r}
assessment2017<-assessment_pct  %>%  filter(Year=="2017")
tax2017<-tax_pct[,1:2]
assessment2017<-assessment2017 %>% left_join(tax2017, by = c("Municipalities"))
names(assessment2017)[9]<-paste("tax")

assessment2018<-assessment_pct  %>%  filter(Year=="2018")
tax2018<-tax_pct[,c(1,3)]
assessment2018<-assessment2018 %>% left_join(tax2018, by = c("Municipalities"))
names(assessment2018)[9]<-paste("tax")

assessment2019<-assessment_pct  %>%  filter(Year=="2019")
tax2019<-tax_pct[,c(1,4)]
assessment2019<-assessment2019 %>% left_join(tax2019, by = c("Municipalities"))
names(assessment2019)[9]<-paste("tax")

pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
write.csv(pct_final, here("data","data_final.csv"))

# all pct data aggregated
#head(pct_final)

```

2019 tax is missing, we imputed 2019 tax using previous values
```{r}
tax_final <- read.csv(here("data","tax_final.csv"))
#head(tax_final)

# tax_final is the budget
tax_modified <- tax_final  %>% mutate(X2019=tax_final$X2018*(1+(((tax_final$X2017-tax_final$X2016)/tax_final$X2016 + (tax_final$X2018-tax_final$X2017)/tax_final$X2017)/2)))

# tax_modified contains all the budget values from 2016 to 2019
#head(tax_modified)
```

global variables
```{r}
# all pct aggregated
pct_final
# all budget values
tax_modified
```

check match between budget and assessTotal*mill_rate
```{r}
# get predicted tax income for all tax classes and each municipality
assessment_aggregate[,10] <- assessment_aggregate$assessTotal*assessment_aggregate$rate/100
names(assessment_aggregate)[10] <- paste("predTaxIncome")

dat_pred_income <- assessment_aggregate %>% select(Year,AddressAssessorMunicipalityDesc,TaxClassCode,predTaxIncome)

dat_pred_income <- aggregate(dat_pred_income$predTaxIncome,by=list(Year=dat_pred_income$Year,Municipalities=dat_pred_income$AddressAssessorMunicipalityDesc),FUN=sum)
names(dat_pred_income)[3] <- paste("predTaxIncome")
head(dat_pred_income)
```

2. plots of mill rate v.s assessments for each tax class
```{r}
# mill rate vesus year for each municipality and tax class
dat_rate <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, rate)
#head(dat_rate)

dat_rate_1 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
dat_rate_5 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
dat_rate_6 <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6")

# plot assessed value vesus year for each municipality and tax class
dat_assess <- assessment_aggregate %>% select(Year, AddressAssessorMunicipalityDesc, TaxClassCode, assessTotal)

dat_assess_1 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="1") 
dat_assess_5 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="5") 
dat_assess_6 <- dat_assess %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list) %>% filter(TaxClassCode=="6")

# plot budgets vesus year
dat_tax_long <- gather(tax_modified, Year, TaxAmount, X2016:X2019, factor_key=TRUE)

# Vancouver/West Vancouver are special cases for assessment, we omit them for taxClass 1 and 6
rate_class1 <- dat_rate_1 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

assess_class1 <- dat_assess_1 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 1 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

multiplot(rate_class1 + theme(legend.position = "none"), assess_class1 + theme(legend.position = "none"))

rate_class5 <- dat_rate_5 %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

assess_class5 <- dat_assess_5 %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 5") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

multiplot(rate_class5 + theme(legend.position = "none"),assess_class5 + theme(legend.position = "none"))

rate_class6 <- dat_rate_6 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=rate,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("mill rate for taxClass 6 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

assess_class6 <- dat_assess_6 %>% filter(AddressAssessorMunicipalityDesc != "Vancouver" & AddressAssessorMunicipalityDesc != "West Vancouver") %>% ggplot(aes(x=Year,y=assessTotal,group=AddressAssessorMunicipalityDesc,color=AddressAssessorMunicipalityDesc)) + geom_point() + ggtitle("assessTotal for taxClass 6 and all municipalities") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

multiplot(rate_class6 + theme(legend.position = "none"),assess_class6 + theme(legend.position = "none"))

budget_plot <- dat_tax_long %>% filter(Municipalities != "Vancouver" & Municipalities != "West Vancouver") %>% ggplot(aes(x=Year,y=TaxAmount,group=Municipalities,color=Municipalities)) + geom_point() + ggtitle("Tax v.s Year") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")
budget_plot
```

3. plots of mill rate pct change v.s assessment pct change for each tax class
```{r}
pct_rate1 <- pct_final %>% 
  filter(Municipalities %in% municipality.list) %>%
  filter(TaxClassCode == "01") %>%
  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
    geom_point() + ggtitle("taxClass1 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

pct_rate5 <- pct_final %>% 
  filter(Municipalities %in% municipality.list) %>%
  filter(TaxClassCode == "05") %>%
  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
    geom_point() + ggtitle("taxClass5 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

pct_rate6 <- pct_final %>% 
  filter(Municipalities %in% municipality.list) %>%
  filter(TaxClassCode == "06") %>%
  ggplot(aes(x=Year, y=rate_pct, group=Municipalities, color=Municipalities)) +
    geom_point() + ggtitle("taxClass6 pct rate v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

pct_assess1 <- pct_final %>% 
  filter(Municipalities %in% municipality.list) %>%
  filter(TaxClassCode == "01") %>%
  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
    geom_point() + ggtitle("taxClass1 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

pct_assess5 <- pct_final %>% 
  filter(Municipalities %in% municipality.list) %>%
  filter(TaxClassCode == "05") %>%
  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
    geom_point() + ggtitle("taxClass5 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

pct_assess6 <- pct_final %>% 
  filter(Municipalities %in% municipality.list) %>%
  filter(TaxClassCode == "06") %>%
  ggplot(aes(x=Year, y=assessTotal_pct, group=Municipalities, color=Municipalities)) +
    geom_point() + ggtitle("taxClass6 pct assessment v.s year") + theme(legend.position = "none") + geom_smooth(aes(group = 1), size = 0.5, method = "lm", se = FALSE, colour = "black")

multiplot(pct_rate1,pct_assess1)
multiplot(pct_rate5,pct_assess5)
multiplot(pct_rate6,pct_assess6)
```

4. pct change of mill rate and assessment for each municipality
```{r}
# for each municipality
tax.list <- c("01","05","06")
for (i in 1:21){
  if (municipality.list[i] != "Lions Bay") {
    plot_rate_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
    filter(TaxClassCode %in% tax.list) %>% 
    ggplot(aes(x=Year, y=rate_pct, group=TaxClassCode, color=TaxClassCode)) +
      geom_line() + ggtitle(municipality.list[i])

    plot_assess_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
    filter(TaxClassCode %in% tax.list) %>% 
    ggplot(aes(x=Year, y=assessTotal_pct, group=TaxClassCode, color=TaxClassCode)) +
      geom_line() + ggtitle(municipality.list[i])

    plot_tax_pct <- pct_final %>% filter(Municipalities == municipality.list[i]) %>% 
  filter(TaxClassCode %in% tax.list) %>% 
  ggplot(aes(x=Year, y=tax, group=TaxClassCode, color=TaxClassCode)) +
    geom_line() + ggtitle(municipality.list[i])

    multiplot(plot_rate_pct,plot_assess_pct,plot_tax_pct)
  }
}
```

5. box plots of mill rates, grouped by region
```{r}
rate_boxes <- ggplot(dat_rate, aes(x=AddressAssessorMunicipalityDesc,y=rate,fill=TaxClassCode)) + geom_boxplot() + facet_wrap(~TaxClassCode) + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none")
rate_boxes 
```

6. Test correlations between past mill rates and assessed values and government budget.
```{r}
# aggregate tax and mill rate & assessment
colnames(tax_modified) <- c("X","Municipalities","2016","2017","2018","2019")
dat_tax_long <- gather(tax_modified, Year, TaxAmount, "2016":"2019", factor_key=TRUE)

tax_modified1 <- gather(tax_modified, Year, tax, '2016':'2019', factor_key = TRUE)%>%
  rename(AddressAssessorMunicipalityDesc = Municipalities)

assessment_aggregate <- merge(assessment_aggregate, tax_modified1, by = c("AddressAssessorMunicipalityDesc","Year"))
assessment_aggregate <- assessment_aggregate[,-c(3,11)]
names(assessment_aggregate)[10] <- paste("tax")

# correlation between mill rate and budget for each municipality
rate_tax_correlation <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(rate, tax))
rate_tax_correlation
# correlation between mill rate and assessment
rate_assess_correlation2 <- assessment_aggregate %>%
  group_by(AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(cor(assessTotal, rate))
rate_assess_correlation2
```

6. Compute autocorrelations of past mill rates to test the independence of mill rates. 
```{r}
# all white noise
for (i in 1:21) {
  print(municipality.list[i])
  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "1") %>% select(rate)
  if (municipality.list[i] != "Maple Ridge Rural") {
    as.ts(dat_tmp)
    acf(dat_tmp)
  }
}
for (i in 1:21) {
  print(municipality.list[i])
  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "5") %>% select(rate)
  if (municipality.list[i] != "Maple Ridge Rural" & municipality.list[i] != "White Rock" & municipality.list[i] != "Anmore" & municipality.list[i] != "Belcarra" & municipality.list[i] != "Lions Bay") {
    as.ts(dat_tmp)
    acf(dat_tmp)
  }
}
for (i in 1:21) {
  print(municipality.list[i])
  dat_tmp <- dat_rate %>% filter(AddressAssessorMunicipalityDesc %in% municipality.list[i] & TaxClassCode %in% "6") %>% select(rate)
  if (municipality.list[i] != "Maple Ridge Rural") {
    as.ts(dat_tmp)
    acf(dat_tmp)
  }
}
```

model fitting
```{r}
linear1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
summary(linear1)

linear2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=pct_final)
summary(linear2)
```
r^2 is too small. we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400). We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882.

```{r}
test<-pct_final %>% filter(Municipalities!="Langley - City") %>% 
  filter(Municipalities!="Langley - Township")
linear_test<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=test)
summary(linear_test)
```

Ridge Regression!
reference: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r
```{r}
library(glmnet)
library(dummies)
dummy_year<-dummy(test$Year)
dummy_municipal<-dummy(test$Municipalities)
dummy_taxclass<-dummy(test$TaxClassCode)
# build x matrix
x<-cbind(dummy_year,dummy_municipal)
x<-cbind(x,dummy_taxclass)
x<-cbind(x,as.numeric(test$assessTotal_pct))
x<-cbind(x,as.numeric(test$tax))
y<-test$rate_pct
lambdas <- 10^seq(2, -3, by = -.1)
dim(x)
colnames(x)[26]<-paste("assessment")
colnames(x)[27]<-paste("tax")
colnames(x)

cv_ridge <- cv.glmnet(x, y, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_train <- predict(cv_ridge, s = optimal_lambda, newx = x)

# Compute R^2 from true and predicted values
eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquare = R_square
)

}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
eval_results(y, predictions_train, train)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test, test)

```
Ridge regression doesn't seem to work well. R square = 0.1967394.

Lasso 
```{r}
lambdas <- 10^seq(2, -3, by = -.1)

# Setting alpha = 1 implements lasso regression
lasso_reg <- cv.glmnet(x, y, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 5)

# Best 
lambda_best <- lasso_reg$lambda.min;lambda_best

lasso_model <- glmnet(x, y, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
eval_results(y, predictions_train, train)
# R squared = 0.2000649	
```
Lasso doesn't seem to work well, but it is still better than ridge. R squared = 0.2000649.

Elastic net: 
reference: https://daviddalpiaz.github.io/r4sl/elastic-net.html
```{r}
library(caret)
tibble::as_tibble(test)
cv_5 = trainControl(method = "cv", number = 5)
ridge = train(
  rate_pct ~ Year + Municipalities + TaxClassCode + assessTotal_pct + tax, data = test,
  method = "glmnet",
  trControl = cv_5
)
ridge
```
apha = 1 which implies lasso! refer to lasso.