### 2020-04-08 src/rf.Rmd
extend rf

---
title: "random forest"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
```

```{r}
dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
dat$municipality <- as.factor(dat$municipality)

factors_tbl = dat %>% 
  group_by(municipality) %>% 
  count(name="mun_count", sort = TRUE) %>% 
  ungroup() %>% 
  mutate(perc = mun_count/sum(mun_count),
         cum_perc = cumsum(perc)) %>% 
  arrange(desc(mun_count)) %>% 
  mutate(rank = row_number(),
         municipality = fct_reorder(municipality, rank)) %>% 
  mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
  select(municipality, col_municipality)

dat <- dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  group_by(PIC) %>% 
  mutate(next.assess = lead(total.assessment, order_by = year),
         past.mill = lag(mill.rate, order_by = year)) %>%
  arrange(PIC) %>% 
  group_by(municipality, year) %>% 
  mutate(n.prop = n()) %>% 
  arrange(desc(n.prop)) %>% 
  distinct(municipality, .keep_all = T)

train <- dat %>%  filter(test.train == "train") 
test <- dat %>%  filter(test.train == "test")
```

```{r}
rf.mill <- randomForest(
  mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit,
  data=train
)
```

```{r}
predict(rf.mill, test)
```