### 2020-04-10 src/modeling - 550/rf_mill.Rmd
modified mill model rf

---
title: "Random forest mill rates"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(ROCR))
suppressPackageStartupMessages(library(predictmeans))
```

```{r}
dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
dat$municipality <- as.factor(dat$municipality)

factors_tbl = dat %>% 
  group_by(municipality) %>% 
  count(name="mun_count", sort = TRUE) %>% 
  ungroup() %>% 
  mutate(perc = mun_count/sum(mun_count),
         cum_perc = cumsum(perc)) %>% 
  arrange(desc(mun_count)) %>% 
  mutate(rank = row_number(),
         municipality = fct_reorder(municipality, rank)) %>% 
  mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
  select(municipality, col_municipality)

avg_assessment_by_municipality = dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  filter(test.train == "train") %>% 
  group_by(municipality, tax.class, year) %>% 
  mutate(avg_assessment = mean(total.assessment, na.rm=TRUE)) %>% 
  select(municipality, year, tax.class, avg_assessment) %>% 
  distinct(municipality, .keep_all = T)

dat_1 <- dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  left_join(avg_assessment_by_municipality) %>% 
  group_by(PIC) %>% 
  mutate(next.assess = lead(avg_assessment, order_by = year),
         past.mill = lag(mill.rate, order_by = year)) %>%
  arrange(PIC) %>% 
  group_by(municipality, year) %>% 
  mutate(n.prop = n()) %>% 
  arrange(desc(n.prop)) %>% 
  distinct(municipality, .keep_all = T) %>% 
  select(-c(tax, improvement.assessment, land.assessment, total.assessment))
# 
# notwant <- c("CA-BC-003-6500000", "CA-BC-223-00100030000", "CA-BC-227-0000020000", "CA-BC-234-01001006", "CA-BC-301-0000001", "CA-BC-315-00001000","CA-BC-316-010012122009", "CA-BC-336-00001000", "CA-BC-338-1000004884000", "CA-BC-556-00001000", "CA-BC-727-03025100", "CA-BC-757-05141000", "CA-BC-759-000147000", "CA-BC-760-001203000", "CA-BC-765-00002001", "CA-BC-771-00003000")
train <- dat_1 %>%  filter(test.train == "train")
test <- dat_1 %>%  filter(test.train == "test")
```

# Random Forest
```{r}
set.seed(0)

rf.mill <- randomForest(
  mill.rate ~  tax.class + municipality + avg_assessment + past.mill, na.action = na.omit, mtry = 4,
  data=train, ntree=500
)

save(rf.mill, file = "../src/shiny_app/rf.mill.rda")
#Evaluate variable importance
importance(rf.mill)
varImpPlot(rf.mill)

# rf.as <- randomForest(
#   next.assess ~ tax.class + municipality + total.assessment + mill.rate,na.action = na.omit,mtry = 4,
#   data=train
# )
# 
# importance(rf.as)
# varImpPlot(rf.as)
# 
# yhat.bag <- predict(rf.mill,newdata=test)
# plot(yhat.bag, test$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
# abline(0,1)
# 
# yhat.bag1 <- predict(rf.as,newdata=test)
# plot(yhat.bag1, test$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
# abline(0,1)
```