### 2020-04-10 src/modeling - 550/rf.Rmd

---
title: "random forest and lme"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(ROCR))
suppressPackageStartupMessages(library(predictmeans))
```

```{r}
dat <- readr::read_delim("test_train_data.txt", delim = ",", col_types = "ciccdddddc")
train <- readr::read_delim("train_data.txt", delim = ",", col_types = "ciccddddd")
test <- readr::read_delim("test_data.txt", delim = ",", col_types = "ciccddddd")
dat$municipality <- as.factor(dat$municipality)

factors_tbl = dat %>% 
  group_by(municipality) %>% 
  count(name="mun_count", sort = TRUE) %>% 
  ungroup() %>% 
  mutate(perc = mun_count/sum(mun_count),
         cum_perc = cumsum(perc)) %>% 
  arrange(desc(mun_count)) %>% 
  mutate(rank = row_number(),
         municipality = fct_reorder(municipality, rank)) %>% 
  mutate(col_municipality = fct_collapse(municipality, other = levels(municipality)[-c(1:52)])) %>% 
  select(municipality, col_municipality)

dat <- dat %>% 
  left_join(factors_tbl) %>% 
  select(-c(municipality)) %>% 
  rename(municipality = col_municipality) %>% 
  group_by(PIC) %>% 
  mutate(next.assess = lead(total.assessment, order_by = year),
         past.mill = lag(mill.rate, order_by = year)) %>%
  arrange(PIC) %>% 
  group_by(municipality, year) %>% 
  mutate(n.prop = n()) %>% 
  arrange(desc(n.prop)) %>% 
  distinct(municipality, .keep_all = T)

notwant <- c("CA-BC-003-6500000", "CA-BC-223-00100030000", "CA-BC-227-0000020000", "CA-BC-234-01001006", "CA-BC-301-0000001", "CA-BC-315-00001000","CA-BC-316-010012122009", "CA-BC-336-00001000", "CA-BC-338-1000004884000", "CA-BC-556-00001000", "CA-BC-727-03025100", "CA-BC-757-05141000", "CA-BC-759-000147000", "CA-BC-760-001203000", "CA-BC-765-00002001", "CA-BC-771-00003000")
train <- dat %>%  filter(test.train == "train" & !PIC %in% notwant)
test <- dat %>%  filter(test.train == "test")
```

# Random Forest
```{r}
set.seed(0)

rf.mill <- randomForest(
  mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit, mtry = 4,
  data=train, ntree=500
)

save(rf.mill, file = "~/Desktop/git_docs/RealEstate/src/shiny_app/rf.mill.rda")
#Evaluate variable importance
importance(rf.mill)
varImpPlot(rf.mill)

rf.as <- randomForest(
  next.assess ~ tax.class + municipality + total.assessment + mill.rate,na.action = na.omit,mtry = 4,
  data=train
)

importance(rf.as)
varImpPlot(rf.as)

yhat.bag <- predict(rf.mill,newdata=test)
plot(yhat.bag, test$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
abline(0,1)

yhat.bag1 <- predict(rf.as,newdata=test)
plot(yhat.bag1, test$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
abline(0,1)
```

# Linear Mixed Effect Model
```{r}
# both random slope and random intercept
# different rate of change of assessment value and mill rate as well as initial assessment value and mill rate for each municipality

lme.mill <- lmer(mill.rate ~ 1+ (1+year|municipality) + as.factor(year) + tax.class  + total.assessment + past.mill, data = train)

summary(lme.mill)

lme.as <- lmer(next.assess ~ 1+ (1+year|municipality) + as.factor(year) + tax.class + total.assessment + mill.rate, data = train)

summary(lme.as)
```

```{r}
# check assumptions of lme

# Homogeneity of Variance
residplot(lme.mill)
residplot(lme.as)
```

# Linear Regression
```{r}
lm.as <- lm(
  next.assess ~ PIC + tax.class + municipality + total.assessment + mill.rate, na.action = na.omit, data=train)

predict(lm.as, test)
```